External Representation of Spatial and Spatio-Temporal Values

### 2004-04-27 Documents/STFormat.tex
Representation of Time revised

\section{Instant-Values}
There are several representations for an \emph{instant} value.
All Applications using instants must be able to read each of this.

\paragraph{Date}:
A \emph{<Date>} value is represented as:
\begin{quotation}
  (date <day> <month> <year>)
\end{quotation}
where all values are of type integer. The three values must build a valid
date. A \emph{<Date>} is a shortcut for a \emph{<DateTime>} where all
non-date values are zero.

\paragraph{DateTime}:
A \emph{<DateTime>} value is represented as:
\begin{description}
  \item (datetime <day> <month> <year> <hour> <minute> [ <second> [ <millisecond> ]])
\end{description}

All values are integers. The \emph{<second>} and \emph{<millisecond>} are
optionally (indicated by [...]). If it is omitted, zero is assumed.
\emph{<day>}, \emph{<month>}  and \emph{<year>} must build a
valid date. The remaining values must be in the appropriate range.

\paragraph{Instant}:
A \emph{<Instant>} value is represented as a list containing the symbol {\tt instant}
and a real value (see \ref{real_numbers}).
\begin{description}
  \item (instant <real number>)
\end{description}

\paragraph{Conversions between DateTime and Instant}:
The representation as real value can be computed using the following
mapping (the symbol "datetime" is omitted for a better readability):\\
 $ T : \mbox{\emph{<DateTime>}} \rightarrow \R  $\\
 $ T(d\: m\: y\: h\: {min}\: {sec}\: {msec}) :=
      {ToJulian}(d\, m\, y) +
      \frac{ (((h\cdot 60) + min)\cdot 60 + sec)\cdot 1000+ msec}
           { 86400000 } $\\[1em]

The inverse direction can also be computed.
$ DT : \R \rightarrow \mbox{\emph{<DateTime>}}$\\
The single fields of a \emph{<DateTime>} value can be computed by the following C function.
Note that the granularity is restricted to milliseconds after this conversion.
\begin{footnotesize}
\begin{verbatim}
   void toDateTime(const double instant, int& year, int& month, int& day,
                   int& hour, int& minute, int& second, int& millisecond){
      ToGregorian((long)instant, year,month,day);
      double rest = instant - ((double)((long)instant));
      long R = (long) 86400000*rest;
      millisecond = R % 1000;
      R = R  / 1000;
      second = R % 60;
      R = R / 60;
      minute = R % 60;
      R = R / 60;
      hour = R;
   }
\end{verbatim}
\end{footnotesize}

>>>>

\section{Representation of Time}
There are several representations for an \emph{instant} value.
All Applications using instants must be able to read each of this.
A \emph{DateTime} as an algebra type consist only of its value list. 
If \emph{DateTime} is used as subtype, e.g. in an interval, the
type-name has to be included. The typename can be {\tt datetime} or
{\tt instant}. Summarized we have three possibilities for a \emph{DateTime}:
\begin{tt}
  \begin{small}
     \begin{itemize}
       \item (datetime $<$datetimevalue$>$) \\
       \item (instant $<$datetimevalue$>$)  \\
       \item $<$datetimevalue$>$
      \end{itemize}
  \end{small}
\end{tt}

The representations of {\tt $<$datetimevalue$>$} are described
in the next subsections.

\subsection{String Representation}
\label{StringRepOfTime}
A $<$datetimevalue$>$ can be represented as string in the following
format: \\
\begin{tt}
  \begin{small}
      year-month-day[-hour:minute[:second[.millisecond]]]
  \end{small}
\end{tt}\\
The squared brackets indicate optionally parts. All contained values
are integers which must be in the appropriate interval.

\subsection{Real Representation}
A $<$datetimevalue$>$ can just be represented as an real value. 
The conversion beween the formats is described in section 
\ref{TimeConversions}.

\subsection{Julian Representation}
A $<$datetimevalue$>$ in the Julian representation is:\\
\begin{tt}
  \begin{small}
      (day millisecond)
  \end{small}
\end{tt}\\
{\tt day} can be any integer and millisecond is an integer in 
the interval [0,86400000[. {\tt day} represents the difference
to a fixed nulldate and {\tt milliseconds} represents the 
time part of this day.

\subsection{Gregorian Representation}
In the Gregorian Representation a $<$datetimevalue$>$ is written 
down as follow:\\
\begin{tt}
  \begin{small}
      (day month year [ hour minute [ second [ millisecond]]])
  \end{small}
\end{tt}\\
Like in the string representation (see section \ref{StringRepOfTime}),
the squared brackets are not part of the representation but indicates
optionally information. The values of the omitted parts are assumed to be zero.
The triple (day, month, year) must be a valid date. The remaining
values must be in the appropriate interval.

\subsection{Conversions between Representations of Time}
\label{TimeConversions}
The conversion between {\tt real} value and Julian representation
is straightforward:\\
$ JulToReal : int \times int \rightarrow \R$ \\
$ JulToReal(day,millisecond) = day + \frac{millisecond}{86400000}$ \\
The used operators are working on real numbers.\\[1em]
The another direction is computed as follow. Note a possible lost of 
precision.\\
$ RealToJul : \R \rightarrow (int,int)$ \\
$ RealToJul(instant) = (day,milli) $ \\
where:\\
{\small
$ day = \left\{ \begin{array}{ll}
                    int(instant) & \mbox{if } instant > 0 
                                   \mbox{ or } int(instant)=instant\\
                    int(instant)-1 & otherwise
                \end{array} \right. $\\[1em]
$ milli = \left\{ \begin{array}{ll}
                         tmp & \mbox{if } instant > 0 \mbox{ or } int(instant)=instant\\
                         tmp + 86400000 & \mbox{otherwise}
                      \end{array}\right. $ \\[1em]
\hspace*{1cm}where : tmp = int ((instant-int(instant))*86400000+0.5)}\vspace{1cm}

A quadrupel $(hour, min , sec, millisec)$ can be converted into
an single integer representing all milliseconds using the following formula:\\
$ millis : int \times int \times int \times int \rightarrow int $\\
$ millis(h,min,sec,msec) = (((h\cdot 60) + min)\cdot 60 + sec)\cdot 1000+ msec $
\\[1em]
The inverse direction can be computed as follow:\\
$ splitmillis: int \rightarrow (int,int,int,int)$ \\
$ splitmillis(ms) = (h,min,sec,msec)$ \\
where:
\begin{description}
   \item[] $ h =   {ms}/{3600000} \mod 24 $
   \item[] $ min = {ms}/{60000} \mod 60$
   \item[] $ sec = {ms}/{1000} \mod 60$
   \item[] $ msec = ms \mod 1000 $
\end{description}
